---
title: "FINAL REPROCESSING DATASET"
author: "JUDISMA SALI"
date: "2022-12-16"
output:
  pdf_document: default
  html_document: default
---

```{r}
library(tidyverse)
library(bestNormalize)
```

```{r}
radiomicsdf<- read_csv("radiomics_completedata.csv")
View(radiomicsdf)
head(radiomicsdf)
```

# CHECKING FOR NULL AND MISSING VALUES

 The result for checking null and missing values is 0 using *sum(is.n())*. Thus, there is no null and missing values
 
```{r}
sum(is.na(radiomicsdf))
```

```{r,warning=F}
radiomicsdfs=radiomicsdf%>%select_if(is.numeric)
radiomicsdfs=radiomicsdfs[,-1]
radiomicsdf2=apply(radiomicsdfs,2,function(x){ks.test(x,"pnorm")})
```

To have the list of p-value of all variables, the *unlist()* function is used and convert a list to vector.

```{r}
KS_list=unlist(lapply(radiomicsdf2, function(x) x$p.value))
```


```{r}
sum(KS_list<0.05) # not normally distributed

sum(KS_list>0.05) # normally distributed
```
# [1] 428
# [1] 1

#  Thus, we have 428 variables that are not normally distributed and Entropy_cooc.W.ADC is normally distributed.


```{r}
which.max(KS_list)
```

# CHECKING FOR THEN NORMALITY OF THE DATA

Check for normality, if not, normalized the data

Note that we used *Shapiro-Wilk's Test* to check the normality of the dataset

```{r,warning=F}
temdt=radiomicsdf[,c(3,5:length(names(radiomicsdf)))]

temdt=apply(temdt,2,orderNorm)
temdt=lapply(temdt, function(x) x$x.t)
temdt=temdt%>%as.data.frame()
test=apply(temdt,2,shapiro.test)
test=unlist(lapply(test, function(x) x$p.value))
```

```{r,warning=F}
sum(test>0.05) # not normally distributed
```


```{r,warning=F}
sum(test<0.05) # not normally distributed
```

#[1] 0
#[1] 428

# Thus, base on the result above our data is normally distributed.

```{r}
radiomicsdf[,c(3,5:length(names(radiomicsdf)))]=temdt
```

# GETTING THE CORRELATION OF THE WHOLE DATA EXPECT THE CATEGORICAL VARIABLES

```{r}
CorMatrix=cor(radiomicsdf[,-c(1,2)])
heatmap(CorMatrix,Rowv=NA,Colv=NA,scale="none",revC = T)
```

# SPLITTING THE DATA INTO TRAINING (80%) AND TESTING (20%)

```{r}
radiomicsdf$Institution=as.factor(radiomicsdf$Institution)
radiomicsdf$Failure.binary=as.factor(radiomicsdf$Failure.binary)
```

```{r}
splitter <- sample(1:nrow(radiomicsdf), round(nrow(radiomicsdf) * 0.8))
traindt <- radiomicsdf[splitter, ]
testdt  <- radiomicsdf[-splitter, ]